gyro-johannes

### 2023-06-02 paper/ms.tex
update ms

%
% ms specific numbers
%
\newcommand{\nstarwithages}{NNN}

@@

This work is focused on stars surveyed by Kepler for which ages can be
inferred using either stellar rotation, lithium, or both.  These
``age-dateable stars'', $\mathcal{S}$, are a relatively small subset
of the $\approx$160{,}000 Kepler targets.  Rotation periods provide
the majority of the information, since they have been reported for
$\approx$55{,}000 Kepler stars
\citep[e.g.][]{McQuillan_2014,Santos_2021}.  Measurements of the
\ion{Li}{1} 6708\,\AA\ doublet require high-resolution
($R\gtrsim$10{,}000) spectra, and are only available for a few
thousand Kepler objects of interest (KOIs).  We therefore choose to
define our selection function as Kepler stars with measured
rotation periods ($\mathcal{S}$; Section~\ref{subsec:starsel}), and
Kepler planets whose host stars have measured rotation periods
($\mathcal{P}$; Section~\ref{subsec:planetsel}).

\subsection{Star Selection}
\label{subsec:starsel}

@@

%After reviewing the relevant literature and considering the nature of
%the rotation period detection problem,
We ultimately chose to adopt
the results of \citetalias{Santos_2019} and \citetalias{Santos_2021}
as our default rotation periods.  \citetalias{Santos_2021} provides a
detailed comparison against \citetalias{McQuillan_2014}; the brief
summary is that the periods agree for 99.0\% of the 31{,}038 period
detections in common between the two studies.
\citetalias{Santos_2021} classified the 2{,}992 remaining stars from
\citetalias{McQuillan_2014} as ``non-periodic'' based on updated
knowledge of contaminants (e.g.~giant stars and eclipsing binaries)
and visual inspection.  
In addition, 
\citetalias{Santos_2021} report rotation periods for 24{,}182
main-sequence and subgiant FGKM stars that were not reported as
periodic by \citetalias{McQuillan_2014}.
Our interpretation is that these detections are likely more challenging than
those from \citetalias{McQuillan_2014}, perhaps due to a lower
variability amplitude, or a longer rotation period.
A relevant flag for stars missing \citetalias{McQuillan_2014} periods
is included in Table~\ref{tab:stars}.

\subsection{Planet Selection}
\label{subsec:planetsel}

\subsection{The Stellar Samples}
We will refer in this manuscript to a few samples, defined in detail
below.  The gist of each is as follows.
\begin{itemize}
  \item Kepler objects of interest: as defined in the Cumulative KOI
    table at the NASA Exoplanet Archive from 2022 June 27, which is
    distinct from the \texttt{q1\_q17\_dr25} table
    \citep{Thompson_2018}.
\end{itemize}

@@

\paragraph{Temperatures}
\paragraph{Surface Gravities}

\subsection{Rotation Periods}

@@

@@

\end{document}

>>>>

%
% ms specific numbers
%

%%%%%%%%%
% STARS %
%%%%%%%%%
% number of stars monitored by Kepler (quoting Santos19/21)
\newcommand{\nkeplerstars}{$\approx$160{,}000}
% fraction of kepler stars w/ rotation periods
\newcommand{\nstarswithprot}{{\bf $\sim$50{,}000}}
% fraction of stars with rotation periods, with teff/logg in Berger+20
\newcommand{\fracstarswithprotwithbtwenty}{{\bf $\sim$94\%}}
% complement of above
\newcommand{\fracstarswithprotwithoutbtwenty}{{\bf $\sim$6\%}}

%%%%%%%%%%%
% PLANETS %
%%%%%%%%%%%
% number of KOIs in Q1-Q17 DR25 table
\newcommand{\nkois}{{\bf $\sim$8{,}000}} %FIXME precise
% number of KOIs in Q1-Q17 DR25 table that are not known false positives
\newcommand{\nkoisnofp}{{\bf $\sim$4{,}000}} %FIXME precise
% number of KOIs in Q1-Q17 DR25 table that have rotation periods
\newcommand{\nkoiswithprot}{{\bf $\sim$2{,}000}}
% number of KOIs in Q1-Q17 DR25 table that are not known false positives that have rotation periods
\newcommand{\nkoisnofpwithprot}{{\bf $\sim$1{,}000}}

\newcommand{\fracplanetswithprotwithbtwenty}{{\bf $\sim$94\%}}

\newcommand{\nstarwithages}{NNN}

@@

This work is focused on stars surveyed by Kepler for which ages can be
inferred using either stellar rotation, lithium, or both.  These
``age-dateable stars'', $\mathcal{S}$, are a relatively small subset
of the \nkeplerstars\ Kepler targets.  Rotation periods provide
the majority of the information, since they have been reported for
$\approx$55{,}000 Kepler stars
\citep[e.g.][]{McQuillan_2014,Santos_2021}.  Measurements of the
\ion{Li}{1} 6708\,\AA\ doublet require high-resolution
($R\gtrsim$10{,}000) spectra, and are only available for a few
thousand Kepler objects of interest (KOIs).  We therefore choose to
define our selection function as Kepler stars with measured
rotation periods ($\mathcal{S}$; Section~\ref{subsec:starsel}), and
Kepler planets whose host stars have measured rotation periods
($\mathcal{P}$; Section~\ref{subsec:planetsel}).

\subsection{Stars}
\label{subsec:starsel}

@@

%After reviewing the relevant literature and considering the nature of
%the rotation period detection problem,
We chose to adopt the results of \citetalias{Santos_2019} and
\citetalias{Santos_2021} as our default rotation periods.
\citetalias{Santos_2021} provides a detailed comparison against
\citetalias{McQuillan_2014}; the brief summary is that the periods
agree for 99.0\% of the 31{,}038 period detections in common between
the two studies.  \citetalias{Santos_2021} classified the 2{,}992
remaining stars from \citetalias{McQuillan_2014} as ``non-periodic''
based on updated knowledge of contaminants (e.g.~giant stars and
eclipsing binaries) and visual inspection.  In addition,
\citetalias{Santos_2021} report rotation periods for 24{,}182
main-sequence and subgiant FGKM stars that were not reported as
periodic by \citetalias{McQuillan_2014}.  Our interpretation is that
these detections are likely more challenging than those from
\citetalias{McQuillan_2014}, perhaps due to a lower variability
amplitude, or a longer rotation period.  A relevant flag for stars
missing \citetalias{McQuillan_2014} periods is included in
Table~\ref{tab:stars}.

Concatenating the results of \citetalias{Santos_2019} and
\citetalias{Santos_2021} yields \nstarswithprot\ Kepler stars with
rotation periods.
%FIXME

\subsection{Planets}
\label{subsec:planetsel}

We considered two different samples of planets: those in the fully
automated Q1-Q17 DR25 KOI Table \citep{Thompson_2018}, which are
suitable for planet occurrence rate calculations, and also those in
the cumulative KOI table, which includes the best knowledge available
on any given planet candidate, but incorporates human-based vetting
\footnote{See
\url{https://exoplanetarchive.ipac.caltech.edu/docs/PurposeOfKOITable.html}}.
For the main body of this work, we will refer to the former
homogeneous sample, which includes \nkois\ KOIs, \nkoisnofp\ that are
not known false positives;
Appendix~\ref{app:cumulativekoi} highlights some additional young
planets that are omitted in this approach.

%FIXME: add a sentence or two describing the numbers after
%crossmatching against S19/S21 --

Crossmatching the Q1-Q17 DR25 KOI table against our stars with
rotation periods yields \nkoiswithprot\ KOIs with rotation periods, of which \nkoisnofpwithprot\ 
are not known false postives.

@@

% https://gea.esac.esa.int/archive/documentation/GDR3/Data_analysis/chap_cu8par/sec_cu8par_apsis/ssec_cu8par_apsis_gspphot.html

As our default source for stellar temperatures and surface gravities, we adopt
the values from the Gaia-Kepler Stellar Properties Catalog (GKSP;
\citealt{Berger_2020a_catalog}).
The GKSP parameters are available for stars with ``AAA'' 2MASS photometry,
measured parallaxes in Gaia DR2,  and g-band photometry from either the KIC or
the Kepler-INT survey.
The parameters themselves were derived using \texttt{isoclassify} (CITE
Huber2017) to interpolate over the MIST isochrone grids (CITE Choi2016,
Dotter2016), given the SDSS $g$ and 2MASS $K_{\rm s}$ photometry, the Gaia DR2
parallexes, and $V$-band extinction from the (CITE) Green2019 reddening map.
The resulting stellar parameters are available for
\fracstarswithprotwithbtwenty\ of the \nstarswithprot\ Kepler stars with
rotation periods.
For the remaining \fracstarswithprotwithoutbtwenty\ of stars that lack
temperatures and surface gravities, we adopt the values 
from the Gaia DR3 General Stellar Parametrizer from Photometry (GSP-Phot; CITE).
GSP-Phot uses the low-resolution Gaia BP and RP spectra, mean $G$ magnitudes,
and parallaxes, in concert with the PARSEC isochrones (CITE Tang2014, Chen2015,
Pastorelli2020) to infer the stellar effective temperatures, surface gravities,
metallicities, and line-of-sight extinctions.
TODO: plot uncertainties in teff/logg for both parameter sets.

@@

%FIXME: left off here for discussion of HIRES spectra.

@@
